Rapid Cycleway Prioritisation Tool

### 2020-06-11 code/index.Rmd
Fix formatting

The tool identifies priority locations for new cycleways, ranking roads by their ‘cycling potential’ estimated using the Propensity to Cycle Tool ([PCT](https://www.pct.bike/)). This represents the estimated number of cycling trips along this road travelling to or from work or school assuming the Government’s [aim](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) to double cycling by 2025 is met.

@@

The tool also identifies what a ‘cohesive network’ for cycling might look like if we were to consider a wider range of interventions such as closing roads to motorised traffic and creating one-way systems. This has been estimated by joining up roads that have a high cycling potential on some or all their length. The ‘cohesive network’ might help authorities considering area-wide interventions such as pedestrian and cycle zones or modal filters.

@@

Results for each combined and local authorities can be accessed by clicking on the relevant area from the map below or search the authority’s name from the list below.
This opens a separate webpage presenting a map of the areas identifying ‘existing cycleways’, ‘top ranked new cycleways’ and the ‘cohesive network’.
It also presents a table providing various data about the top ranked new cycleways including the cycling potential, length of continuous cycle network (including the existing network), current speed limit and whether the road is wide and/or has spare road lanes.

@@

For example, authorities considering new cycleways may want to review the ‘top ranked new cycleways’ to inform their decision where best to invest. The table below the maps allows users to rank potential new cycleways by various metrics including:

@@

Additionally, authorities considering area-wide measures may wish to consider the data for the ‘cohesive network’. Currently the tool cannot identify specific interventions to promote cycling (e.g. such as pedestrian and cycle zones or modal filters); however, it should give an indication of where such measures might add greatest value.

@@

- Currently the tool has been developed primarily to help authorities prioritise the best locations for new cycleways. While it provides some indication of what a cohesive network might look like, it doesn’t seek to prioritise what is the best type of intervention for a given location.
- The ranking of cycleways may not represent a proxy for which schemes are likely to deliver the best value for money. As a consequence this tool doesn’t remove the need to use the Department for Transport’s Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) to assess whether proposed schemes deliver value for money.
- The best locations for new cycleways may not always be on roads with spare lanes. For example, in some circumstances it may be preferable for new cycleways to be along quiet roads running parallel to major arterial routes.
- The results for the ‘top ranked new cycleways’ doesn’t identify routes under 500m with multiple changes of road name and doesn’t identify short routes (under 500m) linking up existing cycleways. To address this limitation, the tool includes results labelled ‘spare lane(s)’ and ‘estimated width >10m’ which plug these gaps. These layers are hidden by default but can be found by clicking the ‘layer’ button in the top left corner of the map.
- The road network data is from 2017 and so will be slightly out-of-date.
- The model estimates the cycling potential of roads based on the estimated number of trips to and from work and school. Currently it is not possible to estimate cycling potential accounting for all journey purposes including shopping and leisure trips, which would significantly increase the estimated cycling potential.
- The PCT uses data from the 2011 Census and 2011 Cycle to School Census to estimate cycling potential. Hence the estimated cycling potential may represent an underestimate if there has been significant construction active in the area since 2011.
- The tool currently only identifies priority on-road routes for new cycleways. There may also be desirable off-road locations for new cycleways including alongside old railway lines or across open areas.
- Currently the tool allows for new cycleways on trunk roads with speeds of 40mph or more. While in some limited cases this may be desirable (e.g. where trunk roads are causing significant air pollution), in many cases it may not be desirable to reallocate road space from trunk roads to new cycleways. To address this, we have identify which top ranked routes have speed limits of 40mph or more in the tables.

@@

The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
The tool provides a ‘first pass’ assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.

@@

The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool. 
The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government’s aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019]
Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit (see [cyipt.bike](https://www.cyipt.bike/)).

@@

On the basis that plans should build on and extend existing provision, this layer shows existing cycleways.
This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
<!-- This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps). -->
The definition of ‘cycleway’ for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.
Also, because of the variable time lag between new cycleways being built and their appearance on OSM, not all existing cycleways are included in this layer (although most established cycleways in most areas are).
The data was accessed from OpenStreetMap in May 2020 so is up-to-date in many areas.

>>>>

The tool identifies priority locations for new cycleways, ranking roads by their 'cycling potential' estimated using the Propensity to Cycle Tool ([PCT](https://www.pct.bike/)). This represents the estimated number of cycling trips along this road travelling to or from work or school assuming the Government's [aim](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) to double cycling by 2025 is met.

@@

The tool also identifies what a 'cohesive network' for cycling might look like if we were to consider a wider range of interventions such as closing roads to motorised traffic and creating one-way systems. This has been estimated by joining up roads that have a high cycling potential on some or all their length. The 'cohesive network' might help authorities considering area-wide interventions such as pedestrian and cycle zones or modal filters.

@@

Results for each combined and local authorities can be accessed by clicking on the relevant area from the map below or search the authority's name from the list below.
This opens a separate webpage presenting a map of the areas identifying 'existing cycleways', 'top ranked new cycleways' and the 'cohesive network'.
It also presents a table providing various data about the top ranked new cycleways including the cycling potential, length of continuous cycle network (including the existing network), current speed limit and whether the road is wide and/or has spare road lanes.

@@

For example, authorities considering new cycleways may want to review the 'top ranked new cycleways' to inform their decision where best to invest. The table below the maps allows users to rank potential new cycleways by various metrics including:

@@

Additionally, authorities considering area-wide measures may wish to consider the data for the 'cohesive network'. Currently the tool cannot identify specific interventions to promote cycling (e.g. such as pedestrian and cycle zones or modal filters); however, it should give an indication of where such measures might add greatest value.

@@

- Currently the tool has been developed primarily to help authorities prioritise the best locations for new cycleways. While it provides some indication of what a cohesive network might look like, it doesn't seek to prioritise what is the best type of intervention for a given location.
- The ranking of cycleways may not represent a proxy for which schemes are likely to deliver the best value for money. As a consequence this tool doesn't remove the need to use the Department for Transport's Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) to assess whether proposed schemes deliver value for money.
- The best locations for new cycleways may not always be on roads with spare lanes. For example, in some circumstances it may be preferable for new cycleways to be along quiet roads running parallel to major arterial routes.
- The results for the 'top ranked new cycleways' doesn't identify routes under 500m with multiple changes of road name and doesn't identify short routes (under 500m) linking up existing cycleways. To address this limitation, the tool includes results labelled 'spare lane(s)' and 'estimated width >10m' which plug these gaps. These layers are hidden by default but can be found by clicking the 'layer' button in the top left corner of the map.
- The road network data is from 2017 and so will be slightly out-of-date.
- The model estimates the cycling potential of roads based on the estimated number of trips to and from work and school. Currently it is not possible to estimate cycling potential accounting for all journey purposes including shopping and leisure trips, which would significantly increase the estimated cycling potential.
- The PCT uses data from the 2011 Census and 2011 Cycle to School Census to estimate cycling potential. Hence the estimated cycling potential may represent an underestimate if there has been significant construction active in the area since 2011.
- The tool currently only identifies priority on-road routes for new cycleways. There may also be desirable off-road locations for new cycleways including alongside old railway lines or across open areas.
- Currently the tool allows for new cycleways on trunk roads with speeds of 40mph or more. While in some limited cases this may be desirable (e.g. where trunk roads are causing significant air pollution), in many cases it may not be desirable to reallocate road space from trunk roads to new cycleways. To address this, we have identify which top ranked routes have speed limits of 40mph or more in the tables.

@@

The tool provides an evidence base to support investment in cycling infrastructure, in response to new pressures placed on transport and public health systems by the COVID-19 crisis, at local and regional levels.
The rational for the tool is that, in the context of COVID-19, big decisions around where to prioritise investment on the road network need to be made quickly.
The tool provides a 'first pass' assessment of promising locations for new cycleways which have high latent demand for cycling and where road space can be reallocated whilst maintaining two-way traffic flows for motorists.

@@

The cycling potential of each road is based on the Government Target (equality) scenario from the Propensity to Cycle Tool. 
The figures given represent the numbers of one way cycle trips to work and school that could be expected on each road, in a scenario where the [ Government's aim of an overall doubling of cycling trips has been met](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy).
Our figures include data on travel to work from 2011 Census data and travel to school, assuming cycling grows evenly nationwide accounting for distance and hilliness [@lovelace_propensity_2017; @goodman_scenarios_2019]
Data on road widths and number of lanes is derived from the Cycling Infrastructure Prioritisation Toolkit (see [cyipt.bike](https://www.cyipt.bike/)).

@@

On the basis that plans should build on and extend existing provision, this layer shows existing cycleways.
This may be helpful to view in connection with the other layers, because it can reveal where gaps in the existing cycle network could be filled by new infrastructure.
<!-- This layer does not include some on-road cycle lanes or cycleways less than 500m in length, but these can be viewed in one of the basemap options (see below for more information about the basemaps). -->
The definition of 'cycleway' for the purpose of this layer includes shared use paths but excludes cycle lanes painted on roads.
The quality of the cycle paths is not reported in the data and the presence of a cycleway cannot guarantee that there is adequate provision, because some cycleways included in this dataset are of poor quality and/or are on paths shared by pedestrians that have limited capacity for high cycle traffic while observing physical distancing guidelines.
Also, because of the variable time lag between new cycleways being built and their appearance on OSM, not all existing cycleways are included in this layer (although most established cycleways in most areas are).
The data was accessed from OpenStreetMap in May 2020 so is up-to-date in many areas.
